### 2020-06-12 tables/worst-case-power-budget-flat-traverse-sol-tau-1.tex

\begin{table}[h]
  \footnotesize
  \centering
  \caption{Worst case flat traverse Sol power budget, $\tau$\,=\,1 and $\beta$\,=\,\SI{10}{\degree}.}
  \label{tab:worst-case-traverse-sol-power-budget}
  \begin{tabular}{lc|c|c|c|c|}
    \cline{3-6}
     & & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}Iani Chaos\\ $Ls=\SI{81}{\degree}$\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}Ismenius Cavus\\ $Ls=\SI{273}{\degree}$\end{tabular}} \\ \hline
    \multicolumn{1}{|l|}{Mode} & P {[}W{]} & t {[}min{]} & E {[}Wh{]} & t {[}min{]} & E {[}Wh{]} \\ \hline
    \multicolumn{1}{|l|}{Idle - Day} & 29 & 603 & 291 & 464 & 224 \\ \hline
    \multicolumn{1}{|l|}{Communication} & 52 & 35 & 30 & 35 & 30 \\ \hline
    \multicolumn{1}{|l|}{Traverse} & 113 & 4.8 & 9 & 4.8 & 9 \\ \hline
    \multicolumn{1}{|l|}{Science Stop} & 60 & 60 & 60 & 60 & 60 \\ \hline
    \multicolumn{1}{|l|}{Optimal Pose} & 75 & 12 & 13 & 10 & 13 \\ \hline
    \multicolumn{1}{|l|}{Idle - Night} & 20 & 242 & 242 & 866 & 289 \\ \hline
    \multicolumn{1}{|r|}{Total} & 349 & 1440 & 646 & 1440 & 625 \\ \hline
    \multicolumn{1}{|r|}{+20\% Margin} & 419 & - & 775 & - & 750 \\ \hline
  \end{tabular}
\end{table}